Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-22 paper.tex
more progress

\begin{definition}[$s \approx^{*} s$]
Given any two states $s_1$ and $s_2$, we write $s_1 \approx^{*} s_2$
if and only if there exists a state $s_2' \in \text{\ineffCEKD}$ such
that $\judgeDreduceTrans{s_2}{s_2'}$ and $s_1 \approx s_2'$.
\end{definition}

\todo[inline]{Standardize theorem statement -KC}

\begin{lemma}[Weak Bisimulation between \effCEK{C} and \ineffCEKD]
  \label{thm:surely-monoidic-reduce}
  Suppose $C$ is an instance of the Cast ADT
  and the \lazyD{} ADT, $C$ is a monoid,
  and $cast(T,l,T)=id(T)$.
  If
  $s_{1}, s_{1}' \in \text{\effCEK{C}}$,
  $s_{2}, s_{2}' \in \text{\ineffCEKD}$,
  $s_{1} \approx s_{2}$,
  $\judgeSreduce{C}{s_{1}}{s_{1}'}$, and
  $\judgeDreduce{s_{2}}{s_{2}'}$, then
  \todo[inline]{The premise $\judgeDreduce{s_{2}}{s_{2}'}$ is
    unusual for a weak bisimulation. Makes me suspicious.
    That is usually in the conclusion. If this is indeed correct,
    then you'll need to justify/explain it. -JS}
  \[
    \exists s_2'' \in \text{\ineffCEKD}: \;
    \judgeDreduceTrans{s_{2}'}{s_{2}''} \;\text{and}\; s_1' \approx s_2''
  \]
  Or diagrammatically, 
  \[
  \begin{array}{clclc}
    s_{1} & \longrightarrow_{\text{\effCEK{C}}} & s_1'\\
    \rotatebox[origin=c]{90}{$\approx$} 
    & & & \;\;\;\rotatebox[origin=c]{-20}{$\approx$} \\
    s_{2} & 
    \longrightarrow_{\text{\ineffCEKD}} & s_2' &
    \longrightarrow^{*}_{\text{\ineffCEKD}} & s_2'' \\
  \end{array}
  \]
\end{lemma}
\begin{proof}
  By following the transition rule of \ineffCEKD.
  \todo[inline]{What does ``following the transition rules'' mean?
    That's not normal proof jargon. Do you mean case analysis
    on $\judgeSreduce{C}{s_{1}}{s_{1}'}$?
    -JS}
  See the supplementary materials for details.
\end{proof}

@@

\begin{theorem}[\lazyD\ Cast ADT Respects \ineffCEKD]
  \label{thm:surely-lazyD-correct}
  Suppose $C$ is a \lazyD{} Cast ADT.  If
  $\judgetype{\emptyset}{e}{T}$ and $o : T$, then
  \[
  \text{\evalEqv{\ineffCEKD}{\effCEK{C}}}
  \]
\end{theorem}
\begin{proof}
	If we have a Lazy D Cast ADT and it is a monoid, this 
	proof is immediately from Lemma~\ref{thm:surely-monoidic-eval} and 
	Proposition~\ref{thm:surely-lazyD-eqv}. One such representation is to 
	represent casts as a list of triples of type, label, and type, where 
	$id(T)$ is the empty list and $seq(c,c)$ is the list append. Another such 
	representation is the \lazyD\ hypercoercion!
\end{proof}

>>>>

\begin{definition}[$s \approx^{*} s$]
Given any two states $s_1$ and $s_2$, we write $s_1 \approx^{*} s_2$
if and only if there exists a state $s_2' \in \text{\ineffCEKD}$ such
that
\[
s_1 \approx s_2'
\;\text{and}\;
\judgeDreduceTrans{s_2}{s_2'}
\]
\end{definition}

\begin{lemma}[Weak Bisimulation between \effCEK{C} and \ineffCEKD]
	\label{thm:surely-monoidic-reduce}
	Assume $C$ implements Lazy D Cast ADT
	and $C$ is a monoid
	and $cast(T,l,T)=id(T)$,
	$ s_{1} \in \text{\effCEK{C}}$,
	$ s_{2} \in \text{\ineffCEKD}$,
	$ s_{1} \approx s_{2}$,
\begin{enumerate}
	\item If $s_1 = \sOOhalt{o}$ then $s_2 = \sOOhalt{o}$
	\item If $s_2 = \sOOhalt{o}$ then $s_1 = \sOOhalt{o}$
	\item If $\judgeSreduce{C}{s_1}{s_3}$ 
		then $\judgeDreduce{s_2}{s_4}$ for some $s_4$
	\item If $\judgeDreduce{s_2}{s_4}$
		then $\judgeSreduce{C}{s_1}{s_3}$ for some $s_3$
	\item If $\judgeSreduce{C}{s_1}{s_3}$ and $\judgeDreduce{s_2}{s_4}$
		then $s_3 \approx^{*} s_4$
\end{enumerate}
%	Or diagrammatically, 
%	\[\begin{array}{clclc}
%	s_{1} & \longrightarrow_{\text{\effCEK{C}}} & s_1'\\
%	\rotatebox[origin=c]{90}{$\approx$} 
%	& & & \;\;\;\rotatebox[origin=c]{-20}{$\approx$} \\
%	s_{2} & 
%	\longrightarrow_{\text{\ineffCEKD}} & s_2' &
%	\longrightarrow^{*}_{\text{\ineffCEKD}} & s_2'' \\
%	\end{array}\]

%	$ s_{1}, s_{1}' \in \text{\effCEK{C}}$,
%	$ s_{2}, s_{2}' \in \text{\ineffCEKD}$,
%	$ s_{1} \approx s_{2}$,	
%	$ \judgeSreduce{C}{s_{1}}{s_{1}'}$, and
%	$ \judgeDreduce{s_{2}}{s_{2}'}$, then
%	\[ \exists s_2'' \in \text{\ineffCEKD}: \;
%	\judgeDreduceTrans{s_{2}'}{s_{2}''} \;\text{and}\; s_1' \approx s_2'' \]
%	Or diagrammatically, 
%	\[\begin{array}{clclc}
%	s_{1} & \longrightarrow_{\text{\effCEK{C}}} & s_1'\\
%	\rotatebox[origin=c]{90}{$\approx$} 
%	& & & \;\;\;\rotatebox[origin=c]{-20}{$\approx$} \\
%	s_{2} & 
%	\longrightarrow_{\text{\ineffCEKD}} & s_2' &
%	\longrightarrow^{*}_{\text{\ineffCEKD}} & s_2'' \\
%	\end{array}\]

\end{lemma}
\begin{proof}
	Part (1) and (2) follow from the definition of $s \approx s$.
	Part (3) and (4) are true because we have transition rules for every 
	non-halting state.
	Part (5) is more interesting.
	Consider
	\[
	\sOOreturn{\hcvOOcons{v_1}{c_1}{v_2}{c_2}}{\kOOcast{c}{\kOOcar{\kappa}}}
	\approx
	\]
\end{proof}

\todo[inline]{UNDER CONSTRUCTION -KC}

@@

\subsubsection{Correctness of \effCEK{C}}

\begin{theorem}[\effCEK{C} is correct wrt. \ineffCEKD{}]
  \label{thm:surely-lazyD-correct}
  Suppose $C$ is a \lazyD{} Cast ADT.  If
  $\judgetype{\emptyset}{e}{T}$ and $o : T$, then
  \[
  \text{\evalEqv{\effCEK{C}}{\ineffCEKD}}
  \]
\end{theorem}
\begin{proof}
	If we have a Lazy D Cast ADT and it is a monoid, this 
	proof is immediately from Lemma~\ref{thm:surely-monoidic-eval} and 
	Proposition~\ref{thm:surely-lazyD-eqv}. One such representation is to 
	represent casts as a list of triples of type, label, and type, where 
	$id(T)$ is the empty list and $seq(c,c)$ is the list append. Another such 
	representation is the \lazyD\ hypercoercion!
\end{proof}
